---
title: "analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
library(DBI)
library(RSQLite)
library(dplyr)
library(tidyr)
library(gt)

con <- dbConnect(SQLite(), "inst/extdata/jazz.sqlite")
```

## 1. Study Design & Demographics

In this example, I need to get the general information from study_table

```{r}
study_description <-dbGetQuery(con, "SELECT study_id, study_design FROM study_table;")
```

And some manipulations from trial_subject_table, and lookup from drug_table

```{r}
drug_counts <- dbGetQuery(con, "SELECT 
                                  distinct study_id, drug_id,
                                  count(*) d_count
                                FROM trial_subject_table 
                                WHERE study_id = '14-002'
                                GROUP BY drug_id;")

drug_names <- dbGetQuery(con, "SELECT 
                                 distinct drug_id, drug_name
                               FROM drug_table;")

# I COULD push these to SQL, but it becomes unreadable. Let's do it if necessary.

study_subjects <- drug_counts %>% 
  inner_join(drug_names, "drug_id") %>% 
  select(study_id, drug_name, d_count)

study_subject_tbl <- study_subjects %>% 
  unite("drug_name", c(drug_name, d_count), sep = " n=") %>% 
  group_by(study_id) %>% 
  summarize(drug_name = paste(drug_name, collapse = "<br>"), .groups = "drop")
```

Person age info

```{r}
mean_age <- dbGetQuery(con, "SELECT AVG(person_age) mean_age
                             FROM person_table
                             WHERE person_id IN (
                               SELECT distinct person_id
                               FROM trial_subject_table
                               WHERE study_id = '14-002'
                             );")

min_age <- dbGetQuery(con, "SELECT MIN(person_age) min_age
                            FROM person_table
                            WHERE person_id IN (
                              SELECT distinct person_id
                              FROM trial_subject_table
                              WHERE study_id = '14-002'
                            );")

max_age <- dbGetQuery(con, "SELECT MAX(person_age) max_age
                            FROM person_table
                            WHERE person_id IN (
                              SELECT distinct person_id
                              FROM trial_subject_table
                              WHERE study_id = '14-002'
                            );")

age_tbl <- tibble(
  study_id = "14-002",
  mean_age = paste0(round(mean_age, digits = 1), " years (", min_age, " to ", max_age, ")")
)
```

Person sex info

```{r}
sex_count <- dbGetQuery(con, "SELECT person_sex, count(*) as n
                              FROM person_table
                              WHERE person_id IN (
                                           SELECT distinct person_id
                                           FROM trial_subject_table
                                           WHERE study_id = '14-002'
                                         )
                              GROUP BY person_sex;")

sex_tbl <- sex_count %>% 
  mutate(n = paste0(round(n / sum(n), digits = 3) * 100, "%")) %>% 
  unite("sex", c(person_sex, n), sep = ": ") %>% 
  summarize(sex = paste(sex, collapse = "<br>")) %>% 
  mutate(study_id = "14-002")
```

```{r}
gt(
  study_description %>% 
    inner_join(study_subject_tbl, "study_id") %>% 
    inner_join(age_tbl, "study_id") %>% 
    inner_join(sex_tbl, "study_id") 
) %>% 
  fmt_markdown(columns = c("drug_name", "sex"))
```

## 2. Efficacy results at week 12

I need the count summary again. I can honestly use the function from content1 again

```{r}
treatment_group <- get_study_subjects("14-002", system.file("extdata/jazz.sqlite", package = "jazzR"))
```

### MWT mean sleep latency

Mean of means of sleep latency by person at week 0

```{r}
msl_by_person_baseline <- 
  dbGetQuery(con, "SELECT trial_subject_id, AVG(sleep_latency) mean_sleep_latency
                   FROM mwt_table
                   WHERE week = 0
                   GROUP BY trial_subject_id;")

subject_drug_id <- dbGetQuery(con, "SELECT distinct t.trial_subject_id, d.drug_name 
                                    FROM trial_subject_table t
                                    
                                    INNER JOIN drug_table d 
                                    ON t.drug_id = d.drug_id
                                    WHERE t.study_id = '14-002';")

msl_by_drug_baseline <- subject_drug_id %>% 
  inner_join(msl_by_person_baseline, "trial_subject_id")

msl_by_drug_baseline_facts <- msl_by_drug_baseline %>% 
  group_by(drug_name) %>% 
  summarize(msl = round(mean(mean_sleep_latency), digits = 2),
            sd = round(sd(mean_sleep_latency), digits = 2),
            .groups = "drop")

msl_by_drug_baseline_output <- msl_by_drug_baseline_facts %>% 
  unite("mean_baseline_value", c(msl, sd), sep = " (") %>% 
  mutate(mean_baseline_value = paste0(mean_baseline_value, ")")) %>% 
  summarize(mean_baseline_value = paste0(mean_baseline_value, collapse = "<br>"))
```

Mean Change from Baseline (Not going to do ANOVA) is difference between baseline and week 12

```{r}
msl_by_person_end <- 
  dbGetQuery(con, "SELECT trial_subject_id, AVG(sleep_latency) mean_sleep_latency
                   FROM mwt_table
                   WHERE week = 12
                   GROUP BY trial_subject_id;")

subject_drug_id <- dbGetQuery(con, "SELECT distinct t.trial_subject_id, d.drug_name 
                                    FROM trial_subject_table t
                                    
                                    INNER JOIN drug_table d 
                                    ON t.drug_id = d.drug_id
                                    WHERE t.study_id = '14-002';")

msl_by_drug_end <- subject_drug_id %>% 
  inner_join(msl_by_person_end, "trial_subject_id")

msl_by_drug_end_facts <- msl_by_drug_end %>% 
  group_by(drug_name) %>% 
  summarize(msl_end = round(mean(mean_sleep_latency), digits = 2),
            sd_end = round(sd(mean_sleep_latency), digits = 2),
            .groups = "drop")

msl_by_drug_end_output <- msl_by_drug_end_facts %>% 
  unite("mean_baseline_value", c(msl_end, sd_end), sep = " (") %>% 
  mutate(mean_baseline_value = paste0(mean_baseline_value, ")")) %>% 
  summarize(mean_baseline_value = paste0(mean_baseline_value, collapse = "<br>"))

#### Comparing baseline to end (just going to do absolute msl change)

change_baseline_tbl <- msl_by_drug_baseline_facts %>% 
  inner_join(msl_by_drug_end_facts, "drug_name") %>% 
  mutate(change_f_baseline = round(msl_end - msl, digits = 2)) %>% 
  select(drug_name, change_f_baseline)

#### Difference from placebo
placebo <- change_baseline_tbl[change_baseline_tbl$drug_name == "Placebo", "change_f_baseline"][[1]]

change_baseline_placebo_output <- change_baseline_tbl %>% 
  mutate(diff_placebo = change_f_baseline - placebo) %>% 
  summarize(change_f_baseline = paste0(change_f_baseline, collapse = "<br>"),
            diff_placebo = paste0(diff_placebo, collapse = "<br>"))

```

```{r}
cbind(
  tibble(measure = "MWT mean sleep latency"),
  treatment_group,
  msl_by_drug_baseline_output,
  change_baseline_placebo_output
)
```

## 3. Efficacy Visualization

```{r}
ess <- dbGetQuery(con, "SELECT * FROM ess_table;")

subject_drug <- dbGetQuery(con,
  "SELECT distinct t.trial_subject_id, d.drug_name
   FROM trial_subject_table t

   INNER JOIN drug_table d
   ON t.drug_id = d.drug_id
   WHERE t.study_id = '14-002';"
)

dbDisconnect(con)

tbl <- subject_drug %>%
  inner_join(ess, "trial_subject_id") %>%
  group_by(drug_name, week) %>%
  summarize(m_ess = round(mean(ess_score), digits = 2),
            sd = round(sd(ess_score), digits = 2),
            .groups = "drop") %>% 
  mutate(lo = m_ess - sd,
         hi = m_ess + sd)

# library(ggplot2)
# ggplot() +
#   geom_point(data = tbl, aes(x = week, y = m_ess, color = drug_name))
#   geom_error(data = tbl, aes(x = week, ymin = lo, ymax = hi))
```

