---
title: "data-generation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{data-generation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Since I don't have access to any data right now, I'll generate some, using the assumptions outlined in the [paper](https://pp.jazzpharma.com/pi/sunosi.ca.PM-en.pdf). The data model is already on [figma](https://www.figma.com/file/Hrx2iUgZalScAjykfQgqT8/jazzR?node-id=0%3A1).

I'm just going to do SUNOSI drug, study 14-002 for now, and will create more if I have time.

# Study Design Tables

The 4 tables in this section sets up the study, before collecting data. 

## 1. Study Table

This table contains information about the study being done, at the trial level.

One row per trial.

```{r}
library(tibble)
library(dplyr)
library(tidyr)
library(DT)
set.seed(22)

study_table <- tibble(
  study_id = "14-002",
  study_design = "Randomized, double-blind, placebo-controlled, parallel-group study conducted in adults with narcolepsy (with or without cataplexy)",
  n_participants = 240,
  testing_drugs = "SUNOSI 75mg, SUNOSI 150mg, SUNOSI 300mg, Placebo"
)

datatable(study_table)
```

## 2. Person Table

This table contains information of any persons that participate in a study. We'll create 500 people aged 18-70, randomly sampled M/F.

```{r}
person_table <- tibble(
  person_id = paste0("person", 1:500),
  person_age = sample(18:70, size = 500, replace = TRUE),
  person_sex = sample(c("M", "F"), size = 500, replace = TRUE)
)

datatable(person_table)
```

## 3. Drug Table

This table contains information of all drugs, at the dose level. We'll create 3 different doses of SUNOSI, and a placebo.

```{r}
drug_table <- tibble(
  drug_id = paste0("drug", 1:4),
  drug_name = c("SUNOSI 75mg", "SUNOSI 150mg", "SUNOSI 300mg", "Placebo")
)

datatable(drug_table)
```

## 4. Trial Subject Table

This table brings together the above 3 tables, by assigning `drugs` to participating `persons` in a given `study`. The number of rows for a given study is equal to the number of participants. We'll use first 240 `persons`, and randomly assign drugs in 1:1:1:1 ratio.

```{r}
trial_subject_table <- tibble(
  trial_subject_id = paste0("trial_subject", 1:240),
  study_id = "14-002",
  person_id = person_table[1:240, ]$person_id,
  drug_id = sample(rep(unique(drug_table$drug_id), 60), size = 240)
)

datatable(trial_subject_table)
```

# Data Collection Tables

The following tables represent the data collected from the trial subjects during/after studies. 

## 5. ESS Table

The Epworth Sleepiness Scale is a validated 8-item questionnaire by which patients rate their perceived likelihood
of falling asleep in usual daily life activities. The total score ranges from 0 to 24, with higher scores
reflecting greater sleepiness.

ESS is collected at week 0, 1, 4, 8, 12. 

I'll only generate the total scores per week randomly.

```{r}
ess_table <- crossing(
  unique(trial_subject_table[, "trial_subject_id"]),
  week = c(0, 1, 4, 8, 12)
) %>% 
  mutate(
    ess_score = sample(0:24, size = 1200, replace = TRUE)
  )

datatable(ess_table)
```

## 6. PGIc Table

The PGIc is a 7-point scale ranging from “very much improved” to “very much worse,” which assesses the patient’s report of change in symptoms and clinical condition relative to the start of the study. The percentage of patients improved on the PGIc at Week 12 included patients who were minimally improved, much improved or very much improved relative to baseline. 

PGIc is collected at week 0, 12. 

For the sake of simplicity, we'll interpret score 1 as very much worse, and 7 as very much improved. And we'll conclude that a patient's PGIc improved, if their score from baseline to week 12 increased by any amount.

```{r}
pgic_table <- crossing(
  unique(trial_subject_table[, "trial_subject_id"]),
  week = c(0, 12)
) %>% 
  mutate(
    pgic_score = sample(1:7, size = 480, replace = TRUE)
  )

datatable(pgic_table)
```

## 7. MWT Table

The MWT measures an individual’s ability to remain awake during the daytime, based on the duration of sleep latency (i.e., time to sleep onset) when placed in a darkened, quiet environment. For the MWT patients were instructed to remain awake for as long as possible during each of five 40-minute test sessions, and sleep latency was determined from the mean time (minutes) patients could remain awake during the first four test sessions.

MWT is taken at week 0, 1, 4, 12. Five times each.

Patients who participated in this study all had sleep latency of less than 25 minutes. (We'll set 1 minute floor)

```{r}
mwt_table <- crossing(
  unique(trial_subject_table[, "trial_subject_id"]),
  week = c(0, 1, 4, 12),
  session = c(1, 2, 3, 4, 5)
) %>% 
  mutate(
    sleep_latency = sample(1:25, size = 4800, replace = TRUE)
  )

datatable(mwt_table)
```

# Database construction

Now that we have all the data we need, let's store them in a database.

```{r eval=FALSE}
library(DBI)
library(RSQLite)

con <- dbConnect(SQLite(), "jazz.sqlite")

dbWriteTable(con, "study_table", study_table)
dbWriteTable(con, "person_table", person_table)
dbWriteTable(con, "drug_table", drug_table)
dbWriteTable(con, "trial_subject_table", trial_subject_table)
dbWriteTable(con, "ess_table", ess_table)
dbWriteTable(con, "pgic_table", pgic_table)
dbWriteTable(con, "mwt_table", mwt_table)

dbDisconnect(con)
```

We'll validate the data in a [unit test](tests/testthat/test-data_validation.R).
